A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-09-12 faculty.tex
updates

While teaching should take about 40\% of your time (equivalent to research) and your salary is mainly for teaching, it is not as nearly as important as research for tenure.  Moreover, for most new faculty, teaching is not as difficult as getting grants or publishing papers, and thus is often taken lightly by new faculty.  However, teaching is still important for your tenure case, and being a good teacher can be very rewarding. Moreover, innovative integration of research and teaching can help you win good students and even NSF grants.

\section{Teaching Load}
New TT faculty often have a lighter teaching load in their first several years or even throughout their tenure track.  This is to give you time to establish your research program.  However, you should expect to teach at least one course per semester.  The teaching load is usually higher at teaching universities, and lower at research universities. At some universities, you might be able to buy out of teaching by getting a grant.

@@

A strategy that I found effective is that the whole class gets a small bonus point if everyone completes the evaluation. The bonus is calculated as the percentage of responses multiplied by a factor (X). So, if X is 0.5 and everyone participates, they get 0.5 points each. If only half respond, it's 0.25 points each. It might seem small, but it can make a difference for borderline grades. Since I don't curve grades, this is their only chance to boost their scores. Many students have benefited from this in my previous classes, which really grabs their attention. Plus, since evaluations happen before finals and grades aren't finalized, everyone's motivated to participate—they don't want to miss out. Students not only do it themselves but also remind their friends. It's been working well for me: Spring '24 I had a 90\% response rate (25 out of 28 students), and Fall '23 was a perfect 100\% (16 out of 16 students). Before, my response rate was about 50\%, which is considered higher than average in the department.

>>>>

While teaching takes about 40\% of your time, it is not as nearly as important as research for tenure.  Moreover, for most new faculty, teaching is not as difficult as getting grants or publishing papers, and thus is often taken lightly by new faculty.  However, teaching is still important for your tenure case, and being a good teacher can be very rewarding. Moreover, innovative integration of research and teaching can help you get good students and even NSF grants.

\section{Teaching Load}
New TT faculty often have a lighter teaching load in their first several years or even throughout their tenure track.  This is to give you time to establish your research program.  However, you should expect to teach at least one course per semester. The teaching load is usually higher at teaching universities, and lower at research universities.

The desirable teaching load for TT at most R1 universities is 1+1 (1 course / semester) during the tenure track. While some univs might give you a lighter load in your first year, most univs expect you to teach 1+1 from your second year. In many cases, 1+1 is only for the first 2 or 3 years and after that it increases to 1+2 (or 2+1). It is rare that a TT in CS at R1 universities would have to teach 2+2 during the tenure track.

In many cases, your contract might list 1+1 for the first X year and 1+2 after until tenure, but the dept unofficial gives TT faculty 1+1 until tenure. This is because the dept chair can reduce the teaching load based on performance. There are also other ways to reduce teaching load such as \emph{buying out} of teaching with a grant, leading a very large grant, or have taught many large classes.

TT faculty will teach both undergrad and grad courses.  Grad courses include seminars, special topics, and core courses.  Most new faculty use seminars and special topics to recruit students for their research group.  Undergrad courses are typically higher level courses in your area of expertise (e.g., 400 level course on Software Testing).  It is rare that you have to teach 100 level courses (e.g., Intro to Programming) because there are just too many students in these courses.

\begin{commentbox}[Vu]

  At UNL, TT faculty is 1+1 for the first 3 years, and 1+2 after that.  At GMU, it is 1+1 for the first 3 years, but I haven't seen any TT faculty having to teach more than that during their tenure track. In addition, GMU grants all TT faculty a semester off from teaching before tenure. Some faculty use this to work on their NSF CAREER proposal while others use it to prepare their tenure application.

\end{commentbox}

@@

\begin{commentbox}[Vu]
A strategy that I found effective is that the whole class gets a small bonus point if everyone completes the evaluation. The bonus is calculated as the percentage of responses multiplied by a factor (X). So, if X is 0.5 and everyone participates, they get 0.5 points each. If only half respond, it's 0.25 points each. It might seem small, but it can make a difference for borderline grades. Since I don't curve grades, this is their only chance to boost their scores. Many students have benefited from this in my previous classes, which really grabs their attention. Plus, since evaluations happen before finals and grades aren't finalized, everyone's motivated to participate—they don't want to miss out. Students not only do it themselves but also remind their friends. It's been working well for me: Spring '24 I had a 90\% response rate (25 out of 28 students), and Fall '23 was a perfect 100\% (16 out of 16 students). Before, my response rate was about 50\%, which is already higher than the average.
\end{commentbox}
